Parsing Postfix log files

### 2008-10-28 doc/logparser-architecture.tex
First pass through corrections for sections 1-3.
Second pass through corrections for sections 1-3.

@@

Decoupling the actions from the framework simplifies both framework and
actions: the framework provides services to the actions, and does not need
to deal with the complications that arise, or the task of reconstructing a
mail's journey through Postfix; actions benefit from having services
provided by the framework, freeing them to concentrate on the task of
accurately reconstructing each mail's journey through Postfix and dealing
with the complications described in \sectionref{additional complications}.

@@

There is some similarity between the parser's design and William Wood's
\ATN{}~\cite{atns, nlpip}, a tool used in Computational Linguistics for
creating grammars to parse or generate sentences.  The resemblance between
\ATN{} and the parser is accidental, but it is obvious that the two
different approaches share an underlying separation of concerns; this
appears to be a natural division of responsibility and functionality.
\Tableref{Similarities between ATN and this architecture} shows the
similarities.

>>>>

\newpage{}

@@

Decoupling the actions from the framework simplifies both framework and
actions: the framework provides services to the actions, and does not need
to deal with the complications that arise, or the task of reconstructing a
mail's journey through Postfix; actions benefit from having services
provided by the framework, freeing them to concentrate on the task of
accurately reconstructing each mail's journey through Postfix and dealing
with the complications described in \sectionref{complications}.

@@

There is some similarity between the parser's design and William Wood's
\ATN{}~\cite{atns,nlpip}, a tool used in Computational Linguistics for
creating grammars to parse or generate sentences.  The resemblance between
\ATN{} and the parser is accidental, but it is obvious that the two
different approaches share an underlying separation of concerns; this
appears to be a natural division of responsibility and functionality.
\Tableref{Similarities between ATN and this architecture} shows the
similarities.

### 2008-10-28 doc/logparser-background.tex
Second pass through corrections for sections 1-3.

This section provides background information helpful in understanding the
remainder of the document.  It begins with a discussion of the motivation
underlying the project, followed by some technical information: the use of
a database as an \API{}\@; a brief introduction to \SMTP{}\@; a longer
introduction to Postfix, concentrating on anti-spam restrictions and policy
servers.  The assumptions made in designing and implementing the parser are
explained, as are the conventions used in this document.

@@

This document and the program it describes are part of a larger project to
optimise a mail server's Postfix-based anti-spam restrictions, generate
statistics and graphs, and provide a platform on which new restrictions on
trial can be evaluated to determine if they are beneficial in the fight
against spam.  The program parses Postfix log files and populates a
database with the data gleaned from those log files, providing a consistent
and simple view of the log files that future tools can utilise.  The
gathered data can be used to optimise current anti-spam measures, to
provide a baseline to test new anti-spam measures against, or to produce
statistics showing how effective those measures are.

@@

If the database supports sub-selects percentages can be obtained for the
top ten restrictions (sample output shown in \tableref{Sample output from
SQL query showing percentages}):

@@

The database populated by this program provides a simple interface to
Postfix log files.  Although the interface is a database schema, it is in
effect quite similar to any other \APIlong{}~(\API{}) provided by shared
code: it insulates both user and provider of the \API{} from changes in the
implementation of the \API{}\@.  The algorithm implemented by the parser
can be improved; support can be added for earlier or later releases of
Postfix; bugs can be fixed or limitations removed from the parser; these
changes will not cause the user to be negatively affected.  Statistics and
graphs can be generated from the database; new restrictions can be tested
and the results inspected; trends in the fight against spam can emerge from
historical data saved in the database; the parser remains the same as the
usage adapts.  Using a database simplifies writing programs that need to
interact with the data in several ways:

@@

    \item \SQL{} is reasonably standard and many people will already be
        familiar with it; for those unfamiliar with it there are lots of
        readily available resources from which to learn (a good
        introduction to \SQL{} can be found at~\cite{sql-for-web-nerds},
        others are~\cite{w3schools-sql-tutorial,sqlcourse.com}).  Although
        every vendor implements a different dialect of \SQL{}, the basics
        are the same everywhere (analogous to the overall similarities and
        minor differences between Irish English, British English, American
        English, and Australian English).  Depending on the database in use
        there may be tools available that reduce or remove the requirement
        to know \SQL{}; several are available~\cite{sqlite-guis} for
        SQLite (the default database used by the parser
        implementation).

@@

A detailed description of \SMTP{} is beyond the scope of this document:
introductory guides can be found at~\cite{smtp-intro-01,smtp-intro-02},
and the definitive references are~\cite{RFC821,RFC2821}.

@@

Postfix is a \MTAlong{}~(\MTA{}) with the following design aims (in order
of importance): security, flexibility of configuration, scalability, and
high performance.  It features extensive, extensible, optional anti-spam
restrictions, allowing an administrator to deploy those restrictions which
they judge suitable for their site's needs, rather than a fixed set chosen
by Postfix's author.  These restrictions can be selectively applied,
combined, and bypassed on a per-client, per-recipient, or per-sender basis,
allowing varying levels of stricture and permissiveness.  Postfix leverages
simple lookup tables to support arbitrarily complicated user-defined
sequences of restrictions and exceptions, with policy
servers\footnote{Policy servers will be explained in \sectionref{policy
servers}.} as the ultimate in flexibility.  Administrators can also supply
their own rejection messages to make it clear to senders why exactly their
mail was rejected.  Unfortunately this flexibility has a cost: complexity
in the log files generated.  Although it is easy to use standard Unix text
processing utilities to determine the fate of an individual email,
following the journey an email takes through Postfix can be quite
difficult.  For most mails the journey is simple and brief, but the
remaining minority can have quite complex journeys (see
\sectionref{additional complications} for details).

@@

\subsection{Conventions used in the document}

@@

The words \textit{connection\/} and \textit{mail\/} are often used
interchangeably in this document; in general the word used was chosen based
on the context it appears in.

>>>>

This section provides background information helpful in understanding the
remainder of the thesis.  It begins with a discussion of the motivation
underlying the project, followed by some technical information: the use of
a database as an \API{}\@; a brief introduction to \SMTP{}\@; a longer
introduction to Postfix, concentrating on anti-spam restrictions and policy
servers.  The assumptions made in designing and implementing the parser are
explained, as are the conventions used in this thesis.

@@

This thesis and the program it describes are part of a larger project to
optimise a mail server's Postfix-based anti-spam restrictions, generate
statistics and graphs, and provide a platform on which new restrictions on
trial can be evaluated to determine if they are beneficial in the fight
against spam.  The program parses Postfix log files and populates a
database with the data gleaned from those log files, providing a consistent
and simple view of the log files that future tools can utilise.  The
gathered data can be used to optimise current anti-spam measures, to
provide a baseline to test new anti-spam measures against, or to produce
statistics showing how effective those measures are.

@@

If the database supports sub-selects (where the results of one query are
used in another), percentages can be obtained for the top ten restrictions
(sample output shown in \tableref{Sample output from SQL query showing
percentages}):

@@

The database populated by this program provides a simple interface to
Postfix log files.  Although the interface is a database schema rather than
a set of functions provided by shared code, it is in effect quite similar
to any other \API{}: it provides a stable interface for users of the
interface, allowing the implementation to be changed without adverse
effects.  As long as the \API{} is maintained, the parser can be improved
to handle additional complications; support can be added for earlier or
later releases of Postfix; bugs can be fixed or limitations removed.

Using a database simplifies writing programs that need to interact with the
data in several ways:

@@

    \item \SQL{} is reasonably standard and many people will already be
        familiar with it; for those unfamiliar with it there are lots of
        readily available resources from which to learn (a good
        introduction to \SQL{} can be found at~\cite{sql-for-web-nerds},
        others are~\cite{w3schools-sql-tutorial,sqlcourse.com}).  Although
        every vendor implements a different dialect of \SQL{}, the basics
        are the same everywhere (analogous to the overall similarities and
        minor differences between Irish English, British English, American
        English, and Australian English).  Depending on the database in use
        there may be tools available that reduce or remove the requirement
        to know \SQL{}; several are available for SQLite
        (\url{http://www.sqlite.org/}), the default database used by the
        parser implementation.

@@

A detailed description of \SMTP{} is beyond the scope of this thesis:
introductory guides can be found at~\cite{smtp-intro-01,smtp-intro-02},
and the definitive references are~\cite{RFC821,RFC2821}.

@@

Postfix is a \MTAlong{}~(\MTA{}) with the following design aims (in order
of importance): security, flexibility of configuration, scalability, and
high performance.  It features extensive optional anti-spam restrictions,
allowing an administrator to deploy those restrictions which they judge
suitable for their site's needs, rather than a fixed set chosen by
Postfix's author.  These restrictions can be selectively applied, combined,
and bypassed on a per-client, per-recipient, or per-sender basis, allowing
varying levels of stricture and permissiveness.  Postfix leverages simple
lookup tables to support arbitrarily complicated user-defined sequences of
restrictions and exceptions; policy servers (\sectionref{policy servers})
provide a simple way to write new restrictions.  Administrators can also
supply their own rejection messages to make it clear to senders exactly why
their mail was rejected.  Unfortunately this flexibility has a cost:
complexity in the log files generated.  Although it is easy to use standard
Unix text processing utilities to determine the fate of an individual
email, following the journey an email takes through Postfix can be quite
difficult.  For most mails the journey is simple and brief, but the
remaining minority can have quite complex journeys (see
\sectionref{complications} for details).

@@

\subsection{Conventions used}

@@

The words \textit{connection\/} and \textit{mail\/} are often used
interchangeably; in general the word used was chosen based
on the context it appears in.

### 2008-10-28 doc/logparser-conclusion.tex
Second pass through corrections for sections 1-3.

The real difficulties arise once the parser is successfully dealing with
90\% of the log lines, as the irregularities and complications previously
explained only begin to become apparent once the vast majority of log lines
have been parsed successfully.  Adding new rules to deal with numerous,
infrequently occurring log lines is a simple task, albeit tiresome (though
alleviated by the tool described in \sectionref{adding new actions}),
whereas dealing with mails where information appears to be missing is much
more grueling.  Trawling through the log files, looking for something out
of the ordinary, possibly hundreds or even thousands of lines away from the
last mention of the queueid in question, is extremely time consuming and
error prone.  Sometimes the task is not to spot the line that is unusual,
but to spot that a line normally present is missing, i.e.\ to realise that
one line amongst hundreds is absent.  In all cases the evidence must be
used to construct a hypothesis to explain the irregularities, and that
hypothesis must then be tested in the parser; if successful, the parser
must be modified to deal with the irregularities, without adversely
affecting the existing functionality.  The complications described in
\sectionref{additional complications} were solved in the order they are
described in, and that order closely resembles the frequency with which
they occur; the most frequently occurring complications dominate the
warning messages produced, and so naturally they are the first
complications to be dealt with.

>>>>

The real difficulties arise once the parser is successfully dealing with
90\% of the log lines, as the irregularities and complications previously
explained only begin to become apparent once the vast majority of log lines
have been parsed successfully.  Adding new rules to deal with numerous,
infrequently occurring log lines is a simple task, albeit tiresome (though
alleviated by the tool described in \sectionref{adding new actions}),
whereas dealing with mails where information appears to be missing is much
more grueling.  Trawling through the log files, looking for something out
of the ordinary, possibly hundreds or even thousands of lines away from the
last mention of the queueid in question, is extremely time consuming and
error prone.  Sometimes the task is not to spot the line that is unusual,
but to spot that a line normally present is missing, i.e.\ to realise that
one line amongst hundreds is absent.  In all cases the evidence must be
used to construct a hypothesis to explain the irregularities, and that
hypothesis must then be tested in the parser; if successful, the parser
must be modified to deal with the irregularities, without adversely
affecting the existing functionality.  The complications described in
\sectionref{complications} were solved in the order they are
described in, and that order closely resembles the frequency with which
they occur; the most frequently occurring complications dominate the
warning messages produced, and so naturally they are the first
complications to be dealt with.

### 2008-10-28 doc/logparser-implementation.tex
Second pass through corrections for sections 1-3.

\Figureref{flow chart image part 1} and \figureref{flow chart image part 2}
show the paths the data representing a mail can take through the parser
algorithm.  The flow chart covers the most common paths only; there are
additional, uncommon paths excluded for the sake of clarity; details of the
deviations can be found in \sectionref{additional complications}.

@@

Several more states and corresponding branches are omitted for the sake of
clarity (these extra states are caused by solving the complications
described in \sectionref{additional complications}).

@@

This section presented the core of the parser, starting with a very high
level view and the initial complications that arose.  A flow chart showing
the paths a mail may take through the nascent simplified algorithm was
provided, followed by an explanation of those paths, and a discussion of
the parser's emergent behaviour --- the data from the log files creates the
paths in the flow chart, they are not specified anywhere in the parser.
The framework which holds the parser together was covered next, after which
came a description of the current actions provided by the parser, and the
algorithm for analysing unparsed log lines to create \regexes{} for new
rules.  Detecting, diagnosing and defeating complications forms the largest
single portion of this document, mirroring the development of the parser,
The complications are described in the order they were overcome, with
subsequent problems affecting fewer mails (often by an order or magnitude),
though the time required to solve problems increased with each successive
problem.

@@

\label{additional complications}
\label{complications}

>>>>

\Figureref{flow chart image part 1} and \figureref{flow chart image part 2}
show the paths the data representing a mail can take through the parser
algorithm.  The flow chart covers the most common paths only; there are
additional, uncommon paths excluded for the sake of clarity; details of the
deviations can be found in \sectionref{complications}.

@@

Several more states and corresponding branches are omitted for the sake of
clarity (these extra states are caused by solving the complications
described in \sectionref{complications}).

@@

This section presented the core of the parser, starting with a very high
level view and the initial complications that arose.  A flow chart showing
the paths a mail may take through the nascent simplified algorithm was
provided, followed by an explanation of those paths, and a discussion of
the parser's emergent behaviour --- the data from the log files creates the
paths in the flow chart, they are not specified anywhere in the parser.
The framework which holds the parser together was covered next, after which
came a description of the current actions provided by the parser, and the
algorithm for analysing unparsed log lines to create \regexes{} for new
rules.  Detecting, diagnosing and defeating complications forms the largest
single portion of this section, mirroring the development of the parser.
The complications are described in the order they were overcome, with
subsequent problems affecting fewer mails (often by an order or magnitude),
though the time required to solve problems increased with each successive
problem.

@@

\label{complications}

### 2008-10-28 doc/logparser-introduction.tex
Second pass through corrections for sections 1-3.

Most mail server administrators will have performed some basic processing
of the log files produced by their mail server at one time or another,
whether it was to debug a problem, explain to a user why their mail is
being rejected, or check if new anti-spam measures are working.  The more
adventurous will have generated statistics to show how successful each of
their anti-spam measures has been in the last week, and possibly even
generated some graphs to clearly illustrate these statistics to management
or users.\footnote{This was the author's first real foray into processing
Postfix log files.}  Very few will have performed in-depth parsing and
analysis of their log files, where the parsing must correlate the log lines
per-connection or per-queueid rather than processing log lines
independently.  One of the barriers to this kind of processing is the
unstructured nature of Postfix log files, where each log line was added on
an ad hoc basis as a requirement was discovered or new functionality was
added.\footnote{A history of all changes made to Postfix is distributed
with the source code, available from \url{http://www.postfix.org/}} Further
complication arises because the set of rejection messages is not fixed: new
messages can be added by the administrator with custom checks; every
\DNSBL{}\footnote{This document is supplied with a glossary, see
\textsection\ref{Glossary}.} returns a different explanatory message;
policy servers may log different messages depending on the characteristics
of the connection; there are many ways in which the log lines may differ
between servers, even within the same organisation --- servers may be
configured differently, or running different versions of Postfix.  This
paper documents the difficult process of parsing Postfix log files,
presenting \PLP{}, a program that parses Postfix log files and places the resulting
data into a database for later use.  The gathered data can then be used to
optimise current anti-spam measures, provide a baseline to test new
anti-spam measures against, or to produce statistics showing how effective
those measures are.  Numerous other uses are possible for such data:
improving server performance by identifying troublesome destinations and
reconfiguring appropriately; identifying regular high volume uses (e.g.\
customer newsletters) and restricting those uses to off-peak times;
detecting virus outbreaks that propagate via email; as a base for billing
customers on a shared server.  Preserving the raw data enables users to
develop a multitude of uses far beyond those conceived of by the author.

@@

@@

Section~\ref{conclusion} contains the document's conclusion, describing the
results of the research, design and implementation of the parser.

The bibliography contains references to the resources used in developing
the algorithm, writing the program, and preparing this document.  Also
listed are some additional resources expected to be helpful in
understanding \SMTP{}, Postfix, anti-spam techniques, or the document.

Appendix~\ref{Glossary} provides a glossary of terms used in the document.

Appendix~\ref{Acronyms} provides a list of acronyms used in the document.

>>>>

Most mail server administrators will have performed some basic processing
of the log files produced by their mail server at one time or another,
whether it was to debug a problem, explain to a user why their mail is
being rejected, or check if new anti-spam measures are working.  The more
adventurous will have generated statistics to show how successful each of
their anti-spam measures has been in the last week, and possibly even
generated some graphs to clearly illustrate these statistics to management
or users.\footnote{This was the author's first real foray into processing
Postfix log files.}  Very few will have performed in-depth parsing and
analysis of their log files, where the parsing must correlate the log lines
per-connection or per-queueid rather than processing log lines
independently.  One of the barriers to this kind of processing is the
unstructured nature of Postfix log files, where each log line was added on
an ad hoc basis as a requirement was discovered or new functionality was
added.\footnote{A history of all changes made to Postfix is distributed
with the source code, available from \url{http://www.postfix.org/}} Further
complication arises because the set of rejection messages is not fixed: new
messages can be added by the administrator with custom checks; every
\DNSBL{}\footnote{This document is supplied with a glossary, see
\textsection\ref{Glossary}.} returns a different explanatory message;
policy servers may log different messages depending on the characteristics
of the connection.  There are many ways in which the log lines may differ
between servers, even within the same organisation: servers may be
configured differently, or running different versions of Postfix.  This
paper documents the difficult process of parsing Postfix log files,
presenting \PLP{}, a program that parses Postfix log files and places the
resulting data into a database for later use.  The gathered data can then
be used to optimise current anti-spam measures, provide a baseline to test
new anti-spam measures against, or to produce statistics showing how
effective those measures are.  Numerous other uses are possible for such
data: improving server performance by identifying troublesome destinations
and reconfiguring appropriately; identifying regular high volume uses
(e.g.\ customer newsletters) and restricting those uses to off-peak times;
detecting virus outbreaks that propagate via email; as a base for billing
customers on a shared server.  Preserving the raw data enables users to
develop a multitude of uses far beyond those conceived of by the author.

@@

XXX THIS ALL NEEDS TO BE CHECKED AND UPDATED\@.

@@

Section~\ref{conclusion} contains the conclusion of the thesis, describing
the results of the research, design and implementation of the parser.

The bibliography contains references to the resources used in developing
the algorithm, writing the program, and preparing this thesis.  Also
listed are some additional resources expected to be helpful in
understanding \SMTP{}, Postfix, anti-spam techniques, or the thesis.

Appendix~\ref{Glossary} provides a glossary of terms used in the thesis.

Appendix~\ref{Acronyms} provides a list of acronyms used in the thesis.

### 2008-10-28 doc/logparser-other-parsers-review.tex
Second pass through corrections for sections 1-3.

\label{other-parsers}

@@

Ten other parsers, starting with the only previously published research in
this area that the author is aware of, have been reviewed as part of the
background research for this project.  None of the parsers reviewed perform
the kind of advanced parsing and correlation required for this project; all
except one (XXX IS THIS ACCURATE\@?) are intended to perform a specific
parsing and reporting task, whereas \PLP{} is designed to be a generic
parser, extracting data and leaving generation of reports from the data to
other programs.  Some parsers save the data extracted to a data store, but
the majority discard all data once they have finished running, making
historical analysis impossible.  The parsers reviewed all produce a report
of varying complexity and detail, whereas this project's parser does not
produce a report at all; that responsibility is deferred to a separate
program, to be developed later.  The parser developed for this project is
designed to enable much more detailed log analysis by providing a stable
platform for subsequent programs to develop upon.

@@

@@

There only appears to be one prior published paper about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of \LMA{} is quite
different from this parser: it attempts to present correlated log data in a
form suitable for a systems administrator to search using the myriad of
standard Unix text processing utilities already available, producing both a
\CSV{} file and either a MySQL or Berkeley DB database.  The decision to
support both \CSV{} and Berkeley DB appears to have been a serious
limitation: \CSV{} is a very simple format where a record is stored in a
single line, with fields separated by a comma or other punctuation symbol.
Problems with \CSV{} files include the need to escape separators in the
data stored, providing multiple values for a field (e.g.\ multiple
recipients), and adding new fields.  There is no mechanism to document the
fields or the separator, unlike \SQL{} databases where every database
includes a schema naming the fields and the type of data they store
(integer, text, timestamp, etc.).

@@

\LMA{} stores the following date: time and date, hostname and \IP{} address
of client, \IP{} of server, sender and recipient addresses, \SMTP{} code,
and size (for accepted mails only).  It omits XXX XXX\@.  Handling of
multiple recipients, \SMTP{} codes or remote servers\footnote{A single mail
may be sent to multiple remote servers if it was addressed to recipients in
different domains, or Postfix needs to try multiple servers for one or more
recipients.} is not explained in the paper; experimental observation shows
that multiple records are added when there are multiple recipients (sadly
the records are not associated or linked in any way), and presumably the
same approach is taken when there are multiple destination servers.

@@

\LMA{} does provide some simple reports: IP-STORY, FROM-STORY, DAILY-EMAIL
and DAILY-REJECT\@.  The first three extract \CSV{} lines for the specified
client \IP{} address, sender address, or date respectively.  DAILY-REJECT
initially failed with an error message from the Perl interpreter; after
making some corrections to the code it worked, and extracted the \CSV{}
lines for the specified day where the \SMTP{} code signifies a rejection.
All of these reports are trivially simple to produce from the \CSV{} file
using the standard Unix tool \texttt{awk}; the most complicated,
DAILY-REJECT, is merely:

@@

    \item Total number and bytes of mails accepted, delivered and rejected.

    \item Total number of sender and recipient addresses and domains.

@@

    \item For received mail: per-domain totals for mails sent, deferred,
        average delay, maximum delay, and bytes delivered.  For received
        mail: per-domain totals for mails received and bytes received.

    \item Number of mails and bytes sent and received for each address.

@@

@@

@@

Splunk aims to index all an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface supports complicated searches, providing statistics and
graphs in real time, a facility not provided by \parsername{} (report
generation has been deferred to a subsequent program).  Saved searches can
be run periodically and the results emailed to a recipient or sent to an
external program for further processing (though maybe without the graphs
and detailed statistics); the author was unable to save searches, though
that may have been due to limitations in the free version.  The database is
not available for use by external programs, whereas \parsername{} provides
the database and leaves it to the user to utilise it without limit or
restriction.  The interface is optimised for interactive rather than
automated queries and it does not appear to be possible to write
independent tools to utilise the Splunk database; there are additional
reports available on \url{http://www.splunkbase.com/}, but the author was
unable to find any: every category was empty, even those that the interface
claimed had reports available.  Many types of reports are bundled with the
software, though most are variations of a bar or pie chart, except bubble
and heatmap graphs.  It is easy to drill down through the graphs to extract
a portion of the data (e.g.\ select the hour with the largest number of
events, then select a particular host, and finally a specific address),
though it is not possible to search on partial words.

The web interface is quite attractive and simple to use when searching, but
as an administrator it seems unnecessarily difficult to perform simple
tasks.  When testing it took roughly 30 minutes to add a single log file to
be indexed (log files must be indexed before searching), with the downside
that the log file was copied into a spool directory, doubling the disk
space usage.  The most suitable test would be to index all the log files in
a particular directory, but after three hours, numerous futile attempts,
and reading all the available documentation the author admitted defeat.
Using the \CLI{} was more successful: invoking the command \newline{}
\tab{} \texttt{splunk find logs }\textit{log-directory\/}\newline{} added
40 of the \numberOFlogFILES{} log files to the queue for indexing.
Repeated attempts enqueued the same 40 log files, without explaining why
the others were excluded.\footnote{The log files appear to have only been
indexed once; presumably Splunk keeps track of the files it has indexed and
discards requests to index files for a second time.  This may or may not be
a useful feature for \parsername{}.} There did not appear to be an option
to order the log files to ensure they would be processed in the order they
were created.  Subsequently the author was successful in adding a single
file at a time using the \CLI{}; a simple loop to add all desired log files
was sufficient to index all files.  Splunk will periodically check all
those files for updates unless they are manually removed from its list;
this may or may not be useful behaviour.  Splunk did not appear to have any
difficulty in parsing the log files, once instructed to do so.  All
searches performed using the indexed data returned reasonable results.

@@

Splunk supports reading compressed log files without any configuration by
the user.  The free version of Splunk limits the volume of data indexed per
day to 500MB, though a trial Enterprise licence is available that allows
indexing of up to 5GB of data per day.  The cheapest licensed version costs
\$5000 (plus \$1000 support), and still limits the volume of data indexed
per day to 500MB\@.  Typical log file sizes for a small scale mail server
are given in \sectionref{parser efficiency}.

>>>>

@@

XXX UPDATE THE BLURBS\@.

Ten other parsers, starting with the only previously published research in
this area that the author is aware of, have been reviewed as part of the
background research for this project.  Each of the parsers was tested with
the \numberOFlogFILES{} test log files described in \sectionref{parser
efficiency}.  None of the parsers reviewed perform the kind of advanced
parsing and correlation required for this project; all except one (XXX IS
THIS ACCURATE\@?) are intended to perform a specific parsing and reporting
task, whereas \PLP{} is designed to be a generic parser, extracting data
and leaving generation of reports from the data to other programs.  Some
parsers save the data extracted to a data store, but the majority discard
all data once they have finished running, making historical analysis
impossible.  The parsers reviewed all produce a report of varying
complexity and detail, whereas this project's parser does not produce a
report at all; that responsibility is deferred to a separate program, to be
developed later.  The parser developed for this project is designed to
enable much more detailed log analysis by providing a stable platform for
subsequent programs to develop upon.

@@

XXX IS THE BDB STRUCTURE DOCUMENTED\@?

XXX DESCRIBE THE \CSV{} FORMAT\@.

XXX DESCRIBE THE MYSQL SCHEMA\@.

IS THE SAME DATA STORED FOR ALL FORMATS\@?

@@

There only appears to be one prior published paper about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of \LMA{} is quite
different from this parser: it attempts to present correlated log data in a
form suitable for a systems administrator to search using the myriad of
standard Unix text processing utilities already available, producing both a
\CSV{} file and either a MySQL (\url{http://www.mysql.com/}) or Berkeley DB
(\url{http://www.oracle.com/database/berkeley-db/index.html}) database.
XXX SHOULD THOSE URLS BE REFERENCES INSTEAD\@?  The decision to support
both \CSV{} and Berkeley DB appears to have been a serious limitation:
\CSV{} is a very simple format where a record is stored in a single line,
with fields separated by a comma or other punctuation symbol.  Problems
with \CSV{} files include the need to escape separators in the data stored,
providing multiple values for a field (e.g.\ multiple recipients), and
adding new fields.  There is no mechanism to document the fields or the
separator, unlike \SQL{} databases where every database includes a schema
naming the fields and the type of data they store (integer, text,
timestamp, etc.).

@@

\LMA{} stores the following date: time and date, client hostname and \IP{}
address, server \IP{} address, sender and recipient addresses, \SMTP{}
code, and size (for accepted mails only).  It omits XXX XXX\@.  Handling of
multiple recipients, \SMTP{} codes or remote servers\footnote{A single mail
may be sent to multiple remote servers if it was addressed to recipients in
different domains, or Postfix needs to try multiple servers for one or more
recipients.} is not explained in the paper; experimental observation shows
that multiple records are added when there are multiple recipients (sadly
the records are not associated or linked in any way), and presumably the
same approach is taken when there are multiple destination servers.

@@

\LMA{} does provide some simple reports: IP-STORY, FROM-STORY, DAILY-EMAIL
and DAILY-REJECT\@.  The first three extract \CSV{} lines for the specified
client \IP{} address, sender address, or date respectively.  DAILY-REJECT
initially failed with an error message (XXX ADD THE ERROR MESSAGE) from the
Perl interpreter; after making some corrections to the code it worked, and
extracted the \CSV{} lines for the specified day where the \SMTP{} code
signifies a rejection.  All of these reports are trivially simple to
produce from the \CSV{} file using the standard Unix tool \texttt{awk}; the
most complicated, DAILY-REJECT, is merely:

@@

    \item Total number of mails accepted, delivered, and rejected.  Total
        size of mails accepted and delivered.  Total number of sender and
        recipient addresses and domains.

@@

    \item For received mail: per-domain totals for mails sent, deferred,
        average delay, maximum delay, and bytes delivered.  For received
        mail: per-domain totals for size and number of mails received.

    \item Number and size of mails sent and received for each address.

@@

XXX CHECK ALL THIS AGAIN\@; GIVE EXAMPLE PRICES\@.

@@

XXX THE BLURB MENTIONS \SQL{} DB\@; CAN THAT BE ACCESSED SEPARATELY, IS THE
FORMAT DOCUMENTED, ETC\@?

@@

XXX THE FIRST PARAGRAPH NEEDS TO BE REWRITTEN\@; PICK POINTS, THEN COMPARE
AND CONTRAST\@.

Splunk aims to index all an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface supports complicated searches, providing statistics and
graphs in real time, a facility not provided by \parsername{} (report
generation has been deferred to a subsequent program).  Saved searches can
be run periodically and the results emailed to a recipient or sent to an
external program for further processing (though maybe without the graphs
and detailed statistics); the author was unable to save searches, though
that may have been due to limitations in the free version.  The database is
not available for use by external programs, whereas \parsername{} provides
the database and leaves it to the user to utilise it without limit or
restriction.  The interface is optimised for interactive rather than
automated queries and it does not appear to be possible to write
independent tools to utilise the Splunk database; there are additional
reports available on \url{http://www.splunkbase.com/}, but the author was
unable to find any: every category was empty, even those that the interface
claimed had reports available.  Many types of reports are bundled with the
software, though most are variations of a bar or pie chart, except bubble
and heatmap graphs.  It is easy to drill down through the graphs to extract
a portion of the data (e.g.\ select the hour with the largest number of
events, then select a particular host, and finally a specific address),
though it is not possible to search on partial words.  All searches
performed using the indexed data returned reasonable results.  The web
interface is quite attractive and simple to use when searching, but as an
administrator it seems unnecessarily difficult to perform simple tasks.

When testing it took roughly 30 minutes to add a single log file to be
indexed (log files must be indexed before searching), with the downside
that the log file was copied into a spool directory, doubling the disk
space usage.  The most suitable test would be to index all the log files in
a particular directory, but after three hours, numerous futile attempts,
and reading all the available documentation the author admitted defeat.
Using the \CLI{} was more successful: invoking the command \newline{}
\tab{} \texttt{splunk find logs }\textit{log-directory\/}\newline{} added
40 of the \numberOFlogFILES{} log files to the queue for indexing.
Repeated attempts enqueued the same 40 log files, without explaining why
the others were excluded.\footnote{The log files appear to have been
indexed once only; presumably Splunk keeps track of the files it has
indexed and discards requests to index files for a second time.  This may
or may not be a useful feature for \parsername{}.} There did not appear to
be an option to order the log files to ensure they would be processed in
the order they were created, though this may not be necessary or beneficial
with Splunk.  Subsequently the author was successful in adding a single
file at a time using the \CLI{}; a simple loop to add all desired log files
was sufficient to index all files.  Splunk will periodically check all
those files for updates unless they are manually removed from its list;
this may or may not be useful behaviour.  Splunk did not appear to have any
difficulty in parsing the log files, once instructed to do so.
\parsername{} parses the logs it is instructed to parse, in the order
given; periodic parsing of logs is a task an administrator can easily
achieve with \texttt{cron(8)} and \texttt{logrotate(8)}.

@@

Splunk supports reading compressed log files without any configuration by
the user.  The free version of Splunk limits the volume of data indexed per
day to 500MB, though a trial Enterprise licence is available that allows
indexing of up to 5GB of data per day.  In 2007 the cheapest licensed
version cost \$5000 plus \$1000 support, and limited the volume of data
indexed per day to 500MB\@.  Prices were removed from the Splunk website
during 2008; now Splunk's sales team must be contacted for a quote.
Typical log file sizes for a small scale mail server are given in
\sectionref{parser efficiency}.
